Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-05-21 exp1/iswc2013.tex
typo from Arild
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@578 43ef1190-90fd-4dd7-a769-d0121f02b80e
Johns final comment
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@581 43ef1190-90fd-4dd7-a769-d0121f02b80e

Then, we demonstrate how a fractional factorial experiment can be used
to maintain a structured approach to the problem while reducing the
number of runs the experiment needs, and we discuss the trade-offs
involved. We then continue to discuss techniques that shows our
simplistic experiment to be flawed, yet scientifically better than the
practice of benchmarking. Finally, we outline a road-map to establish
the use of these techniques in SPARQL endpoint evaluations both as a
practical tool and as a scientifically more rigorous methodology.

@@

The main objective of this paper is not to establish that the
optimization is significant, it is merely a simulation. It is to
establish a critical practice of evaluations. It is of critical
importance that we understand why some effects are significant. That
``Implement'' is significant is hardly a surprise, that is what we
worked for. That the main effects ``Union'', ``Optional'' and ``Lang''
are strong is also due to that these are fairly demanding things to
evaluate. Since the delay we inserted affects all kinds of joins, it
is also intuitive that interactions between ``Implement'' and those
that require joins are strong. The strong ``BGPComp:Lang'' and
``Lang:Union'' interactions might be due to that the many more triples
needs to be searched for a language tag in one case. However, the
positive ``Implement:TripleC'' interaction evades such explanations;
in fact, it hints that our optimization may not work as well for
larger databases. This is a clear indication that the separation
between levels for the sizes of the data-set is too small and should
be investigated further. Such reasoning should be applied to all
significant effects.

>>>>

Then, we demonstrate how a fractional factorial experiment can be used
to maintain a structured approach to the problem while reducing the
number of runs the experiment needs, and we discuss the trade-offs
involved. We then continue to discuss techniques that show our
simplistic experiment to be flawed, yet scientifically better than the
practice of benchmarking. Finally, we outline a road-map to establish
the use of these techniques in SPARQL endpoint evaluations both as a
practical tool and as a scientifically more rigorous methodology.

@@

The main objective of this paper is not to establish that the
optimization is significant, it is merely a simulation. It is to
establish a critical practice of evaluations. It is of critical
importance that we understand why some effects are significant. That
``Implement'' is significant is hardly a surprise, that is what we
worked for. That the main effects ``Union'', ``Optional'' and ``Lang''
are strong is also due to that these are fairly demanding things to
evaluate. Since the delay we inserted affects all kinds of joins, it
is also intuitive that interactions between ``Implement'' and those
that require joins are strong. The strong ``BGPComp:Lang'' and
``Lang:Union'' interactions might be due to that the many more triples
needs to be searched for a language tag in one case. However, the
positive ``Implement:TripleC'' interaction evades such explanations;
in fact, it hints that our optimization may not work as well for
larger databases. By carefully inspecting table~\ref{tab:pvaluesfull},
we see that we see that whenever ``TripleC'' is at the low lever, the
new implementation is always best, only at the higher level, the old
implementation may be better. This is a clear indication that the
separation between levels for the sizes of the data-set is too small
and should be investigated further. Such reasoning should be applied
to all significant effects.
